Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-09-06 Paper/assa.tex
addressing comments on secs 2, 3 ...

Herein, we use a random vector 
$\vec\xi \in \Omega\in\mathbb{R}^{N_p}$ to parameterize model uncertainties, 
where $N_p$ is the number of uncertain inputs.
In practical computations, the \emph{canonical} variables $\xi_i, i=1,\ldots ,N_p$, are  
mapped to physical ranges meaningful in a given mathematical model. 
As mentioned in the introduction, an active subspace is a low-dimensional subspace
that consists of important directions in a model's input
parameter space~\cite{Constantine:2015}. The effective variability in a model output $f$
due to uncertain inputs is predominantly captured
along these directions. 
The directions constituting the active subspace are the eigenvectors of the positive
semidefinite matrix 
%
\be
\mat{C} = \int_\Omega (\nabla_{\vec{\xi}}f)(\nabla_{\vec{\xi}}f)^\top \mu(d\vec\xi), 
\label{eq:C}
\ee
%
with 
$\mu(d\vec{\xi}) = \pi(\vec{\xi})d\vec{\xi}$, where $\pi(\vec{\xi})$ is the joint probability
distribution of $\vec{\xi}$. Herein, $f$ is assumed to be a square integrable 
function with continuous partial 
derivatives with respect to the input parameters; moreover, we assume the partial derivatives
are square integrable. 
%Hence, it is possible that a given $f(\vec{\xi})$ might not admit an active
%subspace. However, it is of remarkable interest to investigate if one exists
%since the subspace could be exploited to reduce the dimensionality of the
%problem and hence the associated computational effort.
%Here,
%$\nabla_{\vec{\xi}}f$ denotes the gradient vector. 
%with individual components
%being partial derivatives of $f$ with respect to the $i^\text{th}$ input, $\xi_i$. 
Since $\mat{C}$ is symmetric and
positive semidefinite, it admits a spectral decomposition:
%
\be
\mat{C} = \mat{W}\mat{\Lambda}\mat{W}^\top.
\ee
%
Here $\mat{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with the eigenvalues
$\lambda_i$'s sorted in descending order
\[
     \lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_\Np \geq 0,
\] 
and $\mat{W}$ has the (orthonormal) eigenvectors $\vec{w}_1, \ldots, \vec{w}_\Np$ as its columns.
The eigenpairs are partitioned about the $r$th eigenvalue such that
$\lambda_r/\lambda_{r+1}\gg 1$, 
\be
 \mat{W} = [\mat{W}_1~\mat{W}_2],~~\mat{\Lambda} = \begin{bmatrix}\mat{\Lambda}_1 & \\  &
  \mat{\Lambda}_2. 
\end{bmatrix}
\ee
 %
The columns of $\mat{W}_1 = 
\begin{bmatrix} \vec{w}_r \cdots \vec{w}_r\end{bmatrix}$ 
span the dominant eigenspace of $\mat{C}$ and
define the active subspace, and $\mat{\Lambda}_1$ is a diagonal matrix with
the corresponding set of eigenvalues, $\lambda_1, \ldots, \lambda_r$, on its diagonal. 
Once the active subspace
is computed, dimension reduction is accomplished by transforming the parameter
vector $\vec\xi$ into 
$\vec{y} = \mat{W}_1^\top\vec{\xi} \in \R^r$. The elements of $\vec{y}$ are 
referred to as the set of active variables. 
%The number $r$ of active
%variables is equal to the number of dominant eigenvectors.

Consider the function
\[
    G(\vec{y}) = f(\mat{W}_1\vec{y}), \quad \vec{y} \in \R^r.
\]
Following~\cite{Constantine:2015}, we use the approximation 
\[
f(\vec{\xi}) \approx f(\mat{W}_1 \mat{W}_1^T \vec\xi) =  
G(\mat{W}_1^\top \vec{\xi}).
\] 
That is, the model output $f(\vec\xi)$, in the original parameter space,
is approximated by $G(\mat{W}_1^\top \vec{\xi})$ in the active subspace.
%Hence, $f$ is essentially evaluated at the projection of $\vec{\xi}$ on the
%column space of $\mat{W}_1$. 
We could confine uncertainty analysis to the inputs in the
active subspace whose dimension is typically much smaller (in applications that
admit such a subspace) than the dimension of the original input space. To further
expedite uncertainty analysis, one could fit a regression surface to $G$ using the 
following sequence of steps, as outlined in~\cite[chapter 4]{Constantine:2015}. 
\begin{enumerate}
\item Consider a given set of $N$ 
data points, $\big(\vec{\xi}_i, f(\vec{\xi}_i)\big)$, $i = 1, \ldots, N$. 
\item For each $\vec{\xi}_i$, compute $\vec{y}_i = \mat{W}_1^\top\vec{\xi}_i$. Note that
 $G(\vec{y}_i)$ $\approx$ $f(\vec{\xi}_i)$.
\item Use data points $\big(\vec{y}_i, f(\vec\xi_i)\big)$, $i = 1, \ldots, N$, to compute a 
regression surface $\tilde{G}(\vec{y})\approx 
G(\vec{y})$.
\item Overall approximation, $f(\vec{\xi})$ $\approx$ $\tilde{G}(\mat{W}_1^\top\vec{\xi})$.
\end{enumerate}

In practice, the matrix $\mat{C}$ defined in~\eqref{eq:C} is 
approximated using pseudo-random sampling techniques such as Monte Carlo or
Latin hypercube sampling:
%The integral in~\eqref{eq:C} is replaced with a
%summation as follows:
 %
 \be
 \mat{C}\approx \hat{\mat{C}} = \frac{1}{N}\sum\limits_{i=1}^{N} 
 (\nabla_{\vec{\xi}}f(\vec{\xi}_i))(\nabla_{\vec{\xi}}f(\vec{\xi}_i))^\top
 = \hat{\mat{W}}\hat{\mat{\Lambda}}\hat{\mat{W}}^\top
\label{eq:chat}
 \ee
 %
Clearly, the computational effort associated with constructing the matrix
$\hat{\mat{C}}$ scales with the number of samples, $N$. Hence, an iterative
computational approach is adopted in this work to gradually increase  
$N$ until the dominant eigenpairs are approximated
with sufficient accuracy; see Section~\ref{sec:method}.

@@

The total Sobol' index ($T_i(f)$) of a model output, $f(\vec\xi)$ quantifies
the total contribution of the input, $\xi_i$ to the variance of the
output~\cite{Sobol:2001}. Mathematically, this can be expressed as follows:
%
\be
T_i(f) = 1 - 
\frac{\V[\mathbb{E}(f|\vec{\xi}_{\sim i})]}{\V(f)},
\label{eq:total}
\ee
%
where $\vec{\xi}_{\sim i}$ is the input parameter vector with the  
$i^\text{th}$ entry removed, and $\V$ denotes the variance. The total Sobol' index accounts
for the contribution of a given input to the variability in the output by itself
as well as due to its interaction or coupling with other inputs. 
Determining accurate estimates of $T_i(f)$ typically involves tens of
thousands of model runs and is therefore prohibitive in the case of
compute-intensive applications. Derivative based 
global sensitivity measures (DGSMs)~\cite{Sobol:2009} provide a means for
approximating informative upper bounds on $T_i(f)$ at a lower cost; see 
also~\cite{Vohra:2018}.

For $f: \Omega \to \R$, we consider the DGSMs,
\[
    \nu_i(f) := \E{\left(\frac{\partial f}{\partial\xi_i}\right)^2} =
                  \int_\Omega 
                  \left(\frac{\partial f}{\partial\xi_i}\right)^2
                  \pi(\vec{\xi})d\vec{\xi}, \quad i = 1, \ldots, \Np.   
\]
Here $\pi$ is the joint PDF of $\vec\xi$. 
Let $\mat{C}$ be as in~\eqref{eq:C}, and consider the spectral decomposition written 
as $\mat{C} = \sum_{k=1}^\Np \lambda_k \vec{w}_k \vec{w}_k^\top$. Herein, we use the notation 
$\ip{\cdot}{\cdot}$ for the Euclidean inner product.
% where
%$\lambda_i$ are the (non-negative) eigenvalues of $\mat{C}$, in descending
%order, and
%$\vec{w}_k$ are the corresponding (orthonormal) eigenvectors. 
The following
result provides a representation of DGSMs in terms of the 
spectral representation of $\mat{C}$: 
\begin{lemma}
We have
$\nu_i(f) = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2$.
\end{lemma}
\begin{proof}
Note that $\nu_i(f) = \vec{e}_i^\top \mat{C} \vec{e}_i$,  
where $\vec{e}_i$ is the $i$th coordinate vector in $\R^\Np$, $i = 1, \ldots, \Np$.
Therefore,
$\nu_i(f) = \vec{e}_i^T \Big(\sum_{k=1}^\Np \lambda_k \vec{w}_k \vec{w}_k^\top\Big) \vec{e}_i
 = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2$. 
\end{proof}
In the case where the eigenvalues decay rapidly to zero, we can obtain
accurate approximations of $\nu_i(f)$ by truncating the summation: 
\[
   \act{f}{i}{r} =  \sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2,
   \quad i = 1, \ldots, \Np, \quad r \leq \Np.
\]
The quantities $\act{f}{i}{r}$ are called activity scores
in~\cite{Diaz:2016,Constantine:2017}.
%The activity scores connect ideas from active subspaces and global sensitivity
%analysis, and can be used to approximate DGSMs.  
The following result, which
can also be found in~\cite{Diaz:2016,Constantine:2017}, quantifies the error in this
approximation. We provide a short proof for completeness. 
\begin{proposition}\label{prp:dgsm_bound} 
For $1 \leq r < \Np$,
\[
0 \leq \nu_i(f) - \act{f}{i}{r} \leq \lambda_{r+1}, \quad i = 1, \ldots, \Np.
\] 
\end{proposition}
\begin{proof}

@@

Combining Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}, shows
an interesting link between the activity scores and total Sobol' indices.
Specifically, by computing the activity scores, we can identify the unimportant
inputs.  
One can attempt to reduce parameter dimension by fixing
unimportant inputs at nominal values. Suppose $\xi_i$ is
deemed unimportant as a result of a global sensitivity analysis (GSA)
based on the activity scores; we want to estimate
the approximation error that occurs once $\xi_i$ 
is fixed at a nominal value.
%Let $\I \subset \{1, \ldots,
%Np\}$ be an index set that indexes the unimportant parameters in the parameter
%vector $\vec{\xi}$.  Letting $|\I|$ denote the number of elements of $\I$, 
%we denote by $\vec{z} \in \R^{|\I|}$ a generic vector of nominal values for
%the unimportant parameters $\vec\xi_\I$, and let $\vec{y}$ be a generic vector of
%important variables, $\vec\xi_{\I^c}$,
%where $\I^c$ denotes the complement of $\I$ in $\{1, \ldots, \Np\}$. 
%For a vector $\xi$, we let $\xin$ denote the vector obtained by removing
%$\xi_i$. 
%
To formalize this process, we proceed as follows.
Let $\vec\xi$ be given and let $z$ be a nominal value for $\xi_i$.  
%we define 
%$\vec{y}^z(\vec\xi)$ as the vector with entries $y^z_j = \xi_j$ for $j \neq i$
%and $y^z_i = z$.
Consider the \emph{reduced} model, 
obtained by fixing $\xi_i$ at a nominal value $z \in \R$: 
\[
\redf{i}{z} = f(\xi_1, \xi_2, \ldots, \xi_{i-1}, z, \xi_{i+1}, \ldots, \xi_\Np),
\] 
and consider the following error indicator:
\[
\mathcal{E}(z) =
\frac{ \int_\Omega \big( f(\vec\xi) - \redf{i}{z}\big)^2 \, \mu(d\vec\xi) }
          {\int_\Omega f(\vec\xi)^2 \, \mu(d\vec\xi)}.
\] 
\begin{theorem}
We have $\Ez{ \mathcal{E}(z)} \leq 2C_i\big(\act{f}{i}{r} + \lambda_{r+1}\big)/{\V(f)}$, 
for $1 \leq r < \Np$.
\end{theorem}
%
%\[
%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
%\]
%we obtain the following useful inequality, which
%links activity scores and Sobol' indices:
%for $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
%\[
%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
%\]
%This was noted in~\cite{Diaz:2016} for the specific case of all $\xi_i$'s being
%iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
\begin{proof} 
Note that, since 
$\int_\Omega f(\vec\xi)^2 \, \mu(d\vec\xi) = \V(f) + 
\left(\int_\Omega f(\vec\xi) \, \mu(d\vec\xi)\right)^2 \geq \V(f)$, we have
\[
\Ez{ \mathcal{E}(z)} \leq \frac{1}{\V(f)} \Ez{ 
\int_\Omega \big( f(\vec\xi) - \redf{i}{z}\big)^2 \, \mu(d\vec\xi)}
= 2 T_i(f), 
\]
where the equality can be shown using arguments similar to the proof of the main result 
in~\cite{SobolTarantolaGatelliEtAl07}. Using this, along with~\eqref{equ:sobol_bound} and
Proposition~\ref{prp:dgsm_bound}, we have 
\[
\Ez{ \mathcal{E}(z)} \leq 
\frac{2C_i}{\V(f)}\nu_i(f)
\leq 
\frac{2C_i}{\V(f)}\big[\act{f}{i}{r} + \lambda_{r+1}\big]. \qedhere
\]
\end{proof}

@@

\begin{remark}
If the random inputs $\xi_i$, $i = 1, \ldots, \Np$, are iid, then 
the $C_i$'s in the definition of the normalized screening metric will cancel and 
\[
    \tilde{\nu}_i(f) = \frac{\nu_i(f)}{\sum_{i=1}^\Np \nu_i(f)} 
  %    = \frac{\vec{e}_i^T \mat{C} \vec{e}_i}{\trace(\mat{C})} 
      = \frac{\sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2}{\sum_{k = 1}^\Np \lambda_k}.
\]
The expression for the denominator follows from Lemma~\ref{lem:sum}(b). 
Also, in the iid case, using Lemma~\ref{lem:sum}(a) we can simplify the normalized activity scores as follows. 
\[
   \actt{f}{i}{r} =  \frac{\act{f}{i}{r}}{\sum_{i=1}^\Np \act{f}{i}{r}} = 
                     \frac{\sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2}
                          {\sum_{k = 1}^r \lambda_k}.
\]
%The expression for the denominator can be seen from Lemma~\ref{lem:sum}(a).

\end{remark}

>>>>

Herein, we use a random vector 
$\vec\xi \in \Omega\in\mathbb{R}^{N_p}$ to parameterize model uncertainties, 
where $N_p$ is the number of uncertain inputs.
In practical computations, the \emph{canonical} variables $\xi_i, i=1,\ldots ,N_p$, are  
mapped to physical ranges meaningful in a given mathematical model. 
As mentioned in the introduction, an active subspace is a low-dimensional subspace
that consists of important directions in a model's input
parameter space~\cite{Constantine:2015}. The effective variability in a model output $f$
due to uncertain inputs is predominantly captured
along these directions. 
The directions constituting the active subspace are the dominant eigenvectors of the positive
semidefinite matrix 
%
\be
\mat{C} = \int_\Omega (\nabla_{\vec{\xi}}f)(\nabla_{\vec{\xi}}f)^\top \mu(d\vec\xi), 
\label{eq:C}
\ee
%
with 
$\mu(d\vec{\xi}) = \pi(\vec{\xi})d\vec{\xi}$, where $\pi(\vec{\xi})$ is the joint probability
distribution of $\vec{\xi}$. Herein, $f$ is assumed to be a square integrable 
function with continuous partial 
derivatives with respect to the input parameters; moreover, we assume the partial derivatives
are square integrable. 
%Hence, it is possible that a given $f(\vec{\xi})$ might not admit an active
%subspace. However, it is of remarkable interest to investigate if one exists
%since the subspace could be exploited to reduce the dimensionality of the
%problem and hence the associated computational effort.
%Here,
%$\nabla_{\vec{\xi}}f$ denotes the gradient vector. 
%with individual components
%being partial derivatives of $f$ with respect to the $i^\text{th}$ input, $\xi_i$. 
Since $\mat{C}$ is symmetric and
positive semidefinite, it admits a spectral decomposition:
%
\be
\mat{C} = \mat{W}\mat{\Lambda}\mat{W}^\top.
\ee
%
Here $\mat{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with the eigenvalues
$\lambda_i$'s sorted in descending order
\[
     \lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_\Np \geq 0,
\] 
and $\mat{W}$ has the (orthonormal) eigenvectors $\vec{w}_1, \ldots, \vec{w}_\Np$ as its columns.
The eigenpairs are partitioned about the $r$th eigenvalue such that
$\lambda_r/\lambda_{r+1}\gg 1$, 
\be
 \mat{W} = [\mat{W}_1~\mat{W}_2],~~\mat{\Lambda} = \begin{bmatrix}\mat{\Lambda}_1 & \\  &
  \mat{\Lambda}_2. 
\end{bmatrix}
\ee
 %
The columns of $\mat{W}_1 = 
\begin{bmatrix} \vec{w}_r \cdots \vec{w}_r\end{bmatrix}$ 
span the dominant eigenspace of $\mat{C}$ and
define the active subspace, and $\mat{\Lambda}_1$ is a diagonal matrix with
the corresponding set of eigenvalues, $\lambda_1, \ldots, \lambda_r$, on its diagonal. 
Once the active subspace
is computed, dimension reduction is accomplished by transforming the parameter
vector $\vec\xi$ into 
$\vec{y} = \mat{W}_1^\top\vec{\xi} \in \R^r$. The elements of $\vec{y}$ are 
referred to as the set of active variables. 
%The number $r$ of active
%variables is equal to the number of dominant eigenvectors.

Consider the function
\[
    G(\vec{y}) = f(\mat{W}_1\vec{y}), \quad \vec{y} \in \R^r.
\]
Following~\cite{Constantine:2015}, we use the approximation 
\[
f(\vec{\xi}) \approx f(\mat{W}_1 \mat{W}_1^T \vec\xi) =  
G(\mat{W}_1^\top \vec{\xi}).
\] 
That is, the model output $f(\vec\xi)$, in the original parameter space,
is approximated by $G(\mat{W}_1^\top \vec{\xi})$ in the active subspace.
%Hence, $f$ is essentially evaluated at the projection of $\vec{\xi}$ on the
%column space of $\mat{W}_1$. 
We could confine uncertainty analysis to the inputs in the
active subspace whose dimension is typically much smaller (in applications that
admit such a subspace) than the dimension of the original input parameter. To further
expedite uncertainty analysis, one could fit a regression surface to $G$ using the 
following sequence of steps, as outlined in~\cite[chapter 4]{Constantine:2015}. 
\begin{enumerate}
\item Consider a given set of $N$ 
data points, $\big(\vec{\xi}_i, f(\vec{\xi}_i)\big)$, $i = 1, \ldots, N$. 
\item For each $\vec{\xi}_i$, compute $\vec{y}_i = \mat{W}_1^\top\vec{\xi}_i$. Note that
 $G(\vec{y}_i)$ $\approx$ $f(\vec{\xi}_i)$.
\item Use data points $\big(\vec{y}_i, f(\vec\xi_i)\big)$, $i = 1, \ldots, N$, to compute a 
regression surface $\tilde{G}(\vec{y})\approx 
G(\vec{y})$.
\item Overall approximation, $f(\vec{\xi})$ $\approx$ $\tilde{G}(\mat{W}_1^\top\vec{\xi})$.
\end{enumerate}

In practice, the matrix $\mat{C}$ defined in~\eqref{eq:C} is 
approximated using pseudo-random sampling techniques such as Monte Carlo or
Latin hypercube sampling:
%The integral in~\eqref{eq:C} is replaced with a
%summation as follows:
 %
 \be
 \mat{C}\approx \hat{\mat{C}} = \frac{1}{N}\sum\limits_{i=1}^{N} 
 (\nabla_{\vec{\xi}}f(\vec{\xi}_i))(\nabla_{\vec{\xi}}f(\vec{\xi}_i))^\top
 = \hat{\mat{W}}\hat{\mat{\Lambda}}\hat{\mat{W}}^\top
\label{eq:chat}
 \ee
 %
Clearly the computational effort associated with constructing the matrix
$\hat{\mat{C}}$ scales with the number of samples, $N$. Hence, an iterative
computational approach is adopted in this work to gradually increase  
$N$ until the dominant eigenpairs are approximated
with sufficient accuracy; see Section~\ref{sec:method}.

@@

The total Sobol' index ($T_i(f)$) of a model output, $f(\vec\xi)$ quantifies
the total contribution of the input, $\xi_i$ to the variance of the
output~\cite{Sobol:2001}. Mathematically, this can be expressed as follows:
%
\be
T_i(f) = 1 - 
\frac{\V[\mathbb{E}(f|\vec{\xi}_{\sim i})]}{\V(f)},
\label{eq:total}
\ee
%
where $\vec{\xi}_{\sim i}$ is the input parameter vector with the  
$i^\text{th}$ entry removed, and $\V$ denotes the variance. The total Sobol' index accounts
for the contribution of a given input to the variability in the output by itself
as well as due to its interaction or coupling with other inputs. 
Determining accurate estimates of $T_i(f)$ typically involves a large 
number of 
model runs and is therefore can be prohibitive in the case of
compute-intensive applications. Derivative based 
global sensitivity measures (DGSMs)~\cite{Sobol:2009} provide a means for
approximating informative upper bounds on $T_i(f)$ at a lower cost; see 
also~\cite{Vohra:2018}.

For $f: \Omega \to \R$, we consider the DGSMs,
\[
    \nu_i(f) := \E{\left(\frac{\partial f}{\partial\xi_i}\right)^2} =
                  \int_\Omega 
                  \left(\frac{\partial f}{\partial\xi_i}\right)^2
                  \pi(\vec{\xi})d\vec{\xi}, \quad i = 1, \ldots, \Np.   
\]
Here $\pi$ is the joint PDF of $\vec\xi$. 
Let $\mat{C}$ be as in~\eqref{eq:C}, and consider the spectral decomposition written 
as $\mat{C} = \sum_{k=1}^\Np \lambda_k \vec{w}_k \vec{w}_k^\top$. Herein, we use the notation 
$\ip{\cdot}{\cdot}$ for the Euclidean inner product.
% where
%$\lambda_i$ are the (non-negative) eigenvalues of $\mat{C}$, in descending
%order, and
%$\vec{w}_k$ are the corresponding (orthonormal) eigenvectors. 
The following
result provides a representation of DGSMs in terms of the 
spectral representation of $\mat{C}$: 
\begin{lemma}
We have
$\nu_i(f) = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2$.
\end{lemma}
\begin{proof}
Note that $\nu_i(f) = \vec{e}_i^\top \mat{C} \vec{e}_i$,  
where $\vec{e}_i$ is the $i$th coordinate vector in $\R^\Np$, $i = 1, \ldots, \Np$.
Therefore,
$\nu_i(f) = \vec{e}_i^T \Big(\sum_{k=1}^\Np \lambda_k \vec{w}_k \vec{w}_k^\top\Big) \vec{e}_i
 = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2$. 
\end{proof}
In the case where the eigenvalues decay rapidly to zero, we can obtain
accurate approximations of $\nu_i(f)$ by truncating the summation: 
\[
   \act{f}{i}{r} =  \sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2,
   \quad i = 1, \ldots, \Np, \quad r \leq \Np.
\]
The quantities $\act{f}{i}{r}$ are called activity scores
in~\cite{Diaz:2016,Constantine:2017}, where links between GSA measures and
active subspaces is explored.
%The activity scores connect ideas from active subspaces and global sensitivity
%analysis, and can be used to approximate DGSMs.  
The following result, which
can also be found in~\cite{Diaz:2016,Constantine:2017}, quantifies the error in this
approximation. We provide a short proof for completeness. 
\begin{proposition}\label{prp:dgsm_bound} 
For $1 \leq r < \Np$,
\[
0 \leq \nu_i(f) - \act{f}{i}{r} \leq \lambda_{r+1}, \quad i = 1, \ldots, \Np.
\] 
\end{proposition}
\begin{proof}

@@

Combining Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}, shows
an interesting link between the activity scores and total Sobol' indices.
Specifically, by computing the activity scores, we can identify the unimportant
inputs.  
One can attempt to reduce parameter dimension by fixing
unimportant inputs at nominal values.

Suppose activity scores
are used to approximate DGSMs, and suppose
$\xi_i$ is
deemed unimportant as a result, due to a small activity score. 
We want to estimate
the approximation error that occurs once $\xi_i$ is fixed at a nominal value.
%Let $\I \subset \{1, \ldots,
%Np\}$ be an index set that indexes the unimportant parameters in the parameter
%vector $\vec{\xi}$.  Letting $|\I|$ denote the number of elements of $\I$, 
%we denote by $\vec{z} \in \R^{|\I|}$ a generic vector of nominal values for
%the unimportant parameters $\vec\xi_\I$, and let $\vec{y}$ be a generic vector of
%important variables, $\vec\xi_{\I^c}$,
%where $\I^c$ denotes the complement of $\I$ in $\{1, \ldots, \Np\}$. 
%For a vector $\xi$, we let $\xin$ denote the vector obtained by removing
%$\xi_i$. 
%
To formalize this process, we proceed as follows.
Let $\vec\xi$ be given and let $z$ be a nominal value for $\xi_i$.  
%we define 
%$\vec{y}^z(\vec\xi)$ as the vector with entries $y^z_j = \xi_j$ for $j \neq i$
%and $y^z_i = z$.
Consider the \emph{reduced} model, 
obtained by fixing $\xi_i$ at a nominal value $z \in \R$: 
\[
\redf{i}{z} = f(\xi_1, \xi_2, \ldots, \xi_{i-1}, z, \xi_{i+1}, \ldots, \xi_\Np),
\] 
and consider the following relative error indicator:
\[
\mathcal{E}(z) =
\frac{ \int_\Omega \big( f(\vec\xi) - \redf{i}{z}\big)^2 \, \mu(d\vec\xi) }
          {\int_\Omega f(\vec\xi)^2 \, \mu(d\vec\xi)}.
\] 
\begin{theorem}\label{thm:error_estimate}
We have $\Ez{ \mathcal{E}(z)} \leq 2C_i\big(\act{f}{i}{r} + \lambda_{r+1}\big)/{\V(f)}$, 
for $1 \leq r < \Np$.
\end{theorem}
%
%\[
%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
%\]
%we obtain the following useful inequality, which
%links activity scores and Sobol' indices:
%for $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
%\[
%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
%\]
%This was noted in~\cite{Diaz:2016} for the specific case of all $\xi_i$'s being
%iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
\begin{proof} 
Note that, since 
$\int_\Omega f(\vec\xi)^2 \, \mu(d\vec\xi) = \V(f) + 
\left(\int_\Omega f(\vec\xi) \, \mu(d\vec\xi)\right)^2 \geq \V(f)$, we have
\[
\Ez{ \mathcal{E}(z)} \leq \frac{1}{\V(f)} \Ez{ 
\int_\Omega \big( f(\vec\xi) - \redf{i}{z}\big)^2 \, \mu(d\vec\xi)}
= 2 T_i(f), 
\]
where the equality can be shown using arguments similar to the proof of the main result 
in~\cite{SobolTarantolaGatelliEtAl07}. Using this, along with~\eqref{equ:sobol_bound} and
Proposition~\ref{prp:dgsm_bound}, we have 
\[
\Ez{ \mathcal{E}(z)} \leq 
\frac{2C_i}{\V(f)}\nu_i(f)
\leq 
\frac{2C_i}{\V(f)}\big[\act{f}{i}{r} + \lambda_{r+1}\big]. \qedhere
\]
\end{proof}

@@

\begin{remark}
If the random inputs $\xi_i$, $i = 1, \ldots, \Np$, are iid, then 
the $C_i$'s in the definition of the normalized screening metric will cancel and 
\[
    \tilde{\nu}_i(f) = \frac{\nu_i(f)}{\sum_{i=1}^\Np \nu_i(f)} 
  %    = \frac{\vec{e}_i^T \mat{C} \vec{e}_i}{\trace(\mat{C})} 
      = \frac{\sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2}{\sum_{k = 1}^\Np \lambda_k}.
\]
The expression for the denominator follows from Lemma~\ref{lem:sum}(b). 
Also, in the iid case, using Lemma~\ref{lem:sum}(a) we can simplify the normalized activity scores as follows. 
\[
   \actt{f}{i}{r} =  \frac{\act{f}{i}{r}}{\sum_{i=1}^\Np \act{f}{i}{r}} = 
                     \frac{\sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2}
                          {\sum_{k = 1}^r \lambda_k}.
\]
\end{remark}

The significance of the developments in this section are as follows.
Theorem~\ref{thm:error_estimate} provides a theoretical basis for parameter dimension
reduction using activity scores. This is done by providing an estimate of the
error between the reduced model and the original model. If a precise ranking of
parameter importance based on total Sobol' indices is desired, one
can first identify unimportant inputs by computing activity scores, and then
perform a detailed variance based GSA of the remaining model parameters. This approach will 
provide great computational savings as variance based GSA has to done for a small number of
inputs deemed important based on their activity scores. Moreover, the
presented result covers a broad class of input distributions coming from 
Boltzmann family of distributions.
Additionally, the normalized activity scores discussed above provide practical
screening metrics that require only computing the activity scores. This is in
contrast to the bound in Theorem~\ref{thm:error_estimate} that requires the
variance $\V(f)$ of the model output.
